Pottery decor as networks on the Middle Niger

### 2024-03-04 analysis/paper/paper.Rmd
experimenting with clustering sparsified networks

# Load the tidyr package for data tidying and reshaping
library(tidyr)

@@

@@

decor_graph

@@

```{r plot-clusters, fig.cap="Louvain communities within the decor network", echo = FALSE}
# set plot colours for communities
communitycolor <-
  setNames(
    c("tomato", "cornflowerblue", "darkgreen"),
    unique(V(decor_graph)$louvain)
  )

@@

ggraph(decor_graph, x = lon, y = lat) +
  geom_sf(data = mali_adm0, fill = NA) +
  geom_sf(data = rivers, colour = "black") +
  edges_community +
  ggforce::geom_mark_hull(aes(x = lon, y = lat,
                              fill = louvain, colour = louvain),
                          expand = unit(1.5, "mm"),
                          radius = unit(1.5, "mm")) +
  scale_fill_manual(values = communitycolor) +
  scale_colour_manual(values = communitycolor) +
  nodes_community +
  theme_community +
  coord_sf(xlim = c(-7, 2), ylim = c(13, 19), expand = FALSE) +
  theme_graph()
```

@@

```{r cluster1, fig.cap="Internal connections within community 1"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 1) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(size = 3, colour = "cornflowerblue") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

```{r cluster2, fig.cap="Internal connections within community 2"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 2) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(size = 3, colour = "tomato") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

```{r cluster3, fig.cap="Internal connections within community 3"}
decor_graph %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 3) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(size = 3, colour = "darkgreen") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

```{r cluster-comparison, fig.cap="Relative edge weights and centralization by cluster"}
# adding community/crossing to the edges
decor_graph <- decor_graph |>
  add_community_edges()

# mean edge weights
mean_edge_weight <- decor_graph |>
  get.data.frame("edges") |>
  filter(community != "crossing") |>
  group_by(community) |>
  summarise(mean_weight = mean(weight))

# Centralization of the cluster
centralization <- vector("double", length(3))
for (i in 1:3) {
  centralization[[i]] <- decor_graph |>
    # filter community
    activate(nodes) |>
    filter(louvain == i) |>
    # degree centralization
    centr_degree(mode = "all", normalized = TRUE) |>
    as_tibble() |>
    pull(centralization) |>
    head(1)
}

@@

# Graph of centralization of the different communities
g2 <- tibble(centralization) |>
  mutate(community = 1:3) |>
  pivot_longer(!community, names_to = "key", values_to = "centralisation") |>
  ggplot(aes(x = community, y = centralisation)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_grey() +
  theme_bw()

@@

# Calculate centralities for each cluster
centr <- pull_centralities_clust(decor_graph)

@@

#show crossings with a weight over 3
decor_graph |>
  filter(
    community == "crossing",
    weight > 3
  ) |>
  activate(nodes) |>
  filter(!node_is_isolated()) |>
  #plot
  ggraph(layout = "stress") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight),
                      color = louvain), show.legend = FALSE) +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  scale_color_manual(values = communitycolor) +
  theme_graph() +
  theme_community
```

@@

```{r backbone-extraction, fig.cap="Extraction of binary backbones at decreasing significance levels: a) 0.05, b) 0.15, c) 0.18"}
# extract binary backbones at decreasing levels of significance
decor_05 <- extract_backbone(decor_flat, 0.05)
decor_15 <- extract_backbone(decor_flat, 0.15)
decor_18 <- extract_backbone(decor_flat, 0.18)

>>>>

# Load the tidyr package for data tidying and reshaping
library(tidyr)
library(tidyverse)

@@

decor_graph1 <- decor_graph |>
    activate(edges) |> 
   filter(str_detect(name, "roulette") | name %in% c("paint", "straight_mat_impression"))  |> 
  activate(nodes) |>
  mutate(louvain = as.character(group_louvain()))

@@

decor_graph1 <- decor_18 |>
  activate(nodes) |>
  mutate(louvain = as.character(group_louvain()))

decor_graph1 <- decor_graph |>
  activate(nodes) |>
  mutate(louvain = as.data.frame(decor_graph1, "vertices")$louvain)

decor_graph1

decor_graph |>
  activate(edges) |> 
  filter(str_detect(name, "roulette") | name %in% c("paint", "straight_mat_impression"))  |> 
  as.data.frame("edges") |> 
  group_by(name) |> 
  summarise()

decor_graph |>
  activate(edges) |> 
  filter(str_detect(name, "strip_roulette")| str_detect(name, "cord_roulette"))  |> 
  as.data.frame("edges") |> 
  group_by(name) |> 
  summarise()
```

@@

```{r plot-clusters, fig.cap="Louvain communities within the decor network", echo = FALSE}
# set plot colours for communities
communitycolor <-
  setNames(
    c("tomato", "cornflowerblue", "darkgreen", "yellow"),
    unique(V(decor_graph)$louvain)
  )

@@

ggraph(decor_graph1, x = lon, y = lat) +
  geom_sf(data = mali_adm0, fill = NA) +
  geom_sf(data = rivers, colour = "black") +
  edges_community +
  ggforce::geom_mark_hull(aes(x = lon, y = lat,
                              fill = louvain, colour = louvain),
                          expand = unit(1.5, "mm"),
                          radius = unit(1.5, "mm")) +
  scale_fill_manual(values = communitycolor) +
  scale_colour_manual(values = communitycolor) +
  nodes_community +
  theme_community +
  coord_sf(xlim = c(-7, 2), ylim = c(13, 19), expand = FALSE) +
  theme_graph()
```

@@

```{r cluster1, fig.cap="Internal connections within community 1"}
decor_graph1 %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 1) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "cornflowerblue") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

```{r cluster2, fig.cap="Internal connections within community 2"}
decor_graph1 %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 2) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "tomato") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

```{r cluster3, fig.cap="Internal connections within community 3"}
decor_graph1 %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 3) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "darkgreen") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```
```{r cluster3, fig.cap="Internal connections within community 3"}
decor_graph1 %>%
  activate(nodes) %>%
  mutate(deg = centrality_eigen(weights = weight)) %>%
  filter(louvain == 4) %>%
  ggraph(layout = "graphopt") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_color_manual(values = decorcolor[E(decor_graph)$name]) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight)), colour = "yellow") +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  theme_graph()
```

@@

```{r cluster-comparison, fig.cap="Relative edge weights and centralization by cluster"}
# adding community/crossing to the edges
decor_graph1 <- decor_graph1 |>
  add_community_edges()

# mean edge weights
mean_edge_weight <- decor_graph1 |>
  get.data.frame("edges") |>
  filter(community != "crossing") |>
  group_by(community) |>
  summarise(mean_weight = mean(weight))

# Centralization of the cluster
centralization <- vector("double", length(4))
for (i in 1:4) {
  centralization[[i]] <- decor_graph1 |>
    # filter community
    activate(nodes) |>
    filter(louvain == i) |>
    # degree centralization
    centr_degree(mode = "all", normalized = TRUE) |>
    as_tibble() |>
    pull(centralization) |>
    head(1)
}

@@

# Graph of centralization of the different communities
g2 <- tibble(centralization) |>
  mutate(community = 1:4) |>
  pivot_longer(!community, names_to = "key", values_to = "centralisation") |>
  ggplot(aes(x = community, y = centralisation)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_grey() +
  theme_bw()

@@

# Calculate centralities for each cluster
centr <- pull_centralities_clust(decor_graph1)

@@

#show crossings with a weight over 3
decor_graph1 |>
  filter(
    community == "crossing",
    weight > 3
  ) |>
  activate(nodes) |>
  filter(!node_is_isolated()) |>
  #plot
  ggraph(layout = "stress") +
  geom_edge_fan(aes(color = name, edge_width = weight, edge_alpha = weight)) +
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_point(aes(size = centrality_degree(weights = weight),
                      color = louvain), show.legend = FALSE) +
  geom_node_text(aes(label = label), col = "black", repel = TRUE) +
  scale_color_manual(values = communitycolor) +
  theme_graph() +
  theme_community
```

@@

```{r backbone-extraction, fig.cap="Extraction of binary backbones at decreasing significance levels: a) 0.05, b) 0.15, c) 0.18"}
# extract binary backbones at decreasing levels of significance
decor_05 <- extract_backbone(decor_flat, 0.05)
decor_15 <- extract_backbone(decor_flat, 0.15)
decor_18 <- extract_backbone(decor_flat, 0.20) #20 ideale anzahl?
